R-LIME: Rectangular Constraints and Optimization for Local Interpretable Model-agnostic Explanation Methods

### 2024-03-02 paper/paper.tex
Modify detail of images

\section{Proposed Method}
\begin{figure}[tbp]
	\centering
	\begin{subfigure}[t]{0.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{visual-lime}
		\caption{%
			LIME: Locally approximates the decision boundary around the focal point.
		}\label{fig:lime}
	\end{subfigure}%
	\hspace{0.03\textwidth}
	\begin{subfigure}[t]{0.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{visual-anchor}
		\caption{%
			Anchor: Maximizes coverage of a rectangular region
			containing the focal point under accuracy constraints.
		}\label{fig:anchor}
	\end{subfigure}
	\hspace{0.03\textwidth}
	\begin{subfigure}[t]{0.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{visual-rlime}
		\caption{%
			R-LIME:
			Maximizes coverage of a rectangular region containing the focal point
			under lower constraints on the accuracy of the linear classifier.
		}\label{fig:rlime}
	\end{subfigure}
	\caption[Visual comparison of LIME, Anchor and R-LIME]{%
		Visual comparison of LIME, Anchor and R-LIME (our proposed method).
		The solid line represents the rectangular region containing the focal
		point, and the dashed line represents the learned approximation model.
	}
\end{figure}
\subsection{Previous Work}
We specifically focuses on \emph{local} and \emph{model-agnostic} methods.
In this section,
we briefly review existing research on local model-agnostic explanations,
particularly focusing on studies closely related to our proposed method.
\subsubsection[LIME]{%
	LIME (Local Interpretable Model-agnostic Explanations)
	\cite{ribeiro2016why}
}
LIME locally
approximates a trained black-box classifier $f: \mathbb{R}^d \to \{0,1\}$
around a focal point $x \in \mathbb{R}^d$
by a linear classifier $g: \mathbb{R}^d \to \{0,1\}$
(\cref{fig:lime}).
The process involves:
\begin{enumerate}
	\item Obtaining a set of perturbed samples $\mathcal{Z}_p$ around $x$
	      and the set of pseudo-labels $f(\mathcal{Z}_p) = \{f(z) \mid z \in \mathcal{Z}_p\}$.
	\item Learning a linear classifier $g$
	      using $\mathcal{Z}_p$ and $f(\mathcal{Z}_p)$
	      by minimizing the following loss function:
	      \begin{equation}
		      \label{eq:lime_loss}
		      \mathcal{L}(f,g,\pi_x)=\sum_{z\in\mathcal{Z}_p}
		      \pi_x(z){\left(f(z)-g(z)\right)}^2,
	      \end{equation}
	      where $\pi_x(z)$ is a weight function designed to be larger for samples
	      closer to $x$, typically implemented using an exponential kernel.
\end{enumerate}

@@

\ifnum\mode=1
	\subsection{Computational Complexity}
	Post-hoc explanation methods including LIME, Anchor, and R-LIME
	need to query the black-box model multiple times
	to get labels for perturbed samples, which is computationally expensive.
	The number of queries required for LIME is $|\mathcal{Z}_p|$,
	where $|\mathcal{Z}_p|$ is the number of perturbed samples.
	On the other hand, the number of queries required for Anchor and R-LIME
	is $|\mathcal{D}|$,
\fi

@@

		{%
			\def\AB#1{\ifnum#1=0 A\else B\fi}
			\def\mylabel#1{\ifnum#1=0 \label{fig:A-instance}\else \label{fig:B-instance}\fi}
			\ifnum\mode=0
				\renewcommand{\arraystretch}{0.97}
			\else
				\renewcommand{\arraystretch}{1.02}
			\fi
			\begin{figure}[tbp]
				\foreach\a in {0,1}{%
						\centering
						\begin{subfigure}{\textwidth}
							\ifnum\mode=0
								\small
							\fi
							\centering
							\begin{tabular}{p{14em}m{16em}}
								\toprule
								\csvreader[no head, late after line= \\]{%
									\dir/\index{\a}.csv
								}{}{%
								\ifnum\thecsvrow=16 \midrule\fi\csvcoli & \csvcolii
								}
								\bottomrule
							\end{tabular}
							\caption{Instance~\AB{\a}}\mylabel{\a}
							\vspace{15pt}
						\end{subfigure}
					}
				\vspace{-15pt}
				\caption[Two instances sampled from recidivism dataset]{%
					Two instances sampled from recidivism dataset.
				}\label{fig:instance}
			\end{figure}
		}
		{%
			\ifnum\mode=0
				\def\scale{0.3}
				\def\imgwidth{0.45\textwidth}
				\def\hspacebase{\hspace{-0.5em}}
			\else
				\def\scale{0.315}
				\def\imgwidth{0.495\textwidth}
				\def\hspacebase{\hspace{-1.5em}}
			\fi
			\def\vspacebase{\vspace{0.5em}}
			\def\vspacebeforecaption{\vspace{-0.4em}}
			\begin{figure}[p]
				\centering
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/lime-\Asample}
					\vspacebeforecaption
					\caption{LIME}\label{fig:A-lime}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{1.0em}
					\includegraphics[scale=\scale]{\dir/newlime-\Asample-70}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.70$)}\label{fig:A-rlime-70}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/newlime-\Asample-80}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.80$)}\label{fig:A-rlime-80}
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/newlime-\Asample-90}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.90$)}\label{fig:A-rlime-90}
				\end{subfigure}
				\caption[Explanation for Instance A by LIME and R-LIME]{%
					Explanation for Instance A by LIME and R-LIME.
				}\label{fig:A}
			\end{figure}
			\begin{figure}[p]
				\centering
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/lime-\Bsample}
					\vspacebeforecaption
					\caption{LIME}\label{fig:B-lime}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/newlime-\Bsample-70}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.70$)}\label{fig:B-rlime-70}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/newlime-\Bsample-80}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.80$)}\label{fig:B-rlime-80}
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{0em}
					\includegraphics[scale=\scale]{\dir/newlime-\Bsample-90}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.90$)}\label{fig:B-rlime-90}
				\end{subfigure}
				\caption[Explanation for Instance B by LIME and R-LIME]{%
					Explanation for Instance B by LIME and R-LIME.
				}\label{fig:B}
			\end{figure}
		}
	}
\subsubsection{Experimental Results}
The results of the experiment are shown in \cref{fig:A,fig:B}.
The values assigned to each attribute name represent the contribution
(weights of the learned linear classifier)
to the output of the black-box classifier,
normalized such that the absolute sum is 1.
The figures also display the 5 attributes with the highest absolute contribution.

>>>>

\section{Proposed Method}
\begin{figure}[tbp]
	\centering
	\begin{subfigure}[t]{0.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{visual-lime}
		\caption{%
			LIME: Locally approximates the decision boundary around the focal point.
		}\label{fig:lime}
	\end{subfigure}%
	\hspace{0.03\textwidth}
	\begin{subfigure}[t]{0.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{visual-anchor}
		\caption{%
			Anchor: Maximizes coverage of a rectangular region
			containing the focal point under accuracy constraints.
		}\label{fig:anchor}
	\end{subfigure}
	\hspace{0.03\textwidth}
	\begin{subfigure}[t]{0.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{visual-rlime}
		\caption{%
			R-LIME:
			Maximizes coverage of a rectangular region containing the focal point
			under lower constraints on the accuracy of the linear classifier.
		}\label{fig:rlime}
	\end{subfigure}
	\caption[Visual comparison of LIME, Anchor and R-LIME]{%
		Visual comparison of LIME, Anchor and R-LIME (our proposed method).
		The solid line represents the rectangular region containing the focal
		point, and the dashed line represents the learned approximation model.
	}
\end{figure}
\subsection{Previous Work}
We specifically focuses on \emph{local} and \emph{model-agnostic} methods.
In this section,
we briefly review existing research on local model-agnostic explanations,
particularly focusing on studies closely related to our proposed method.
\subsubsection[LIME]{%
	LIME (Local Interpretable Model-agnostic Explanations)
	\cite{ribeiro2016why}
}
LIME locally
approximates a trained black-box classifier $f: \mathbb{R}^m \to \{0,1\}$
around a focal point $x \in \mathbb{R}^m$
by a linear classifier $g: \mathbb{R}^m \to \{0,1\}$
(\cref{fig:lime}).
The process involves:
\begin{enumerate}
	\item Obtaining a set of perturbed samples $\mathcal{Z}_p$ around $x$
	      and the set of pseudo-labels $f(\mathcal{Z}_p) = \{f(z) \mid z \in \mathcal{Z}_p\}$.
	\item Learning a linear classifier $g$
	      using $\mathcal{Z}_p$ and $f(\mathcal{Z}_p)$
	      by minimizing the following loss function:
	      \begin{equation}
		      \label{eq:lime_loss}
		      \mathcal{L}(f,g,\pi_x)=\sum_{z\in\mathcal{Z}_p}
		      \pi_x(z){\left(f(z)-g(z)\right)}^2,
	      \end{equation}
	      where $\pi_x(z)$ is a weight function designed to be larger for samples
	      closer to $x$, typically implemented using an exponential kernel.
\end{enumerate}

@@

\subsection{Computational Complexity}
Post-hoc explanation methods including LIME, Anchor, and R-LIME
need to query the black-box model multiple times
to get labels for perturbed samples, which is computationally expensive.
The number of queries required for LIME is $|\mathcal{Z}_p|$,
where $|\mathcal{Z}_p|$ is the number of perturbed samples.
On the other hand,
the number of queries required for Anchor and R-LIME is
$\mathcal{O}(B\cdot m^2+m^2\cdot\mathcal{O}_{\mathrm{MAB}[B\cdot m,B]})$,
where $B$ is the beam width, $m$ is the number of features, and
$\mathcal{O}_{\mathrm{MAB}[B\cdot m,B]}$ is the computational complexity of
pure exploration
of the multi-armed bandit algorithm with $B\cdot m$ arms and beam width $B$.
$\mathcal{O}_{\mathrm{MAB}[B\cdot m,B]}$ depends on hyperparameters
such as threshold $\tau$, confidence level $1-\delta$ and tolerance $\epsilon$.

@@

		{%
			\def\AB#1{\ifnum#1=0 A\else B\fi}
			\def\mylabel#1{\ifnum#1=0 \label{fig:A-instance}\else \label{fig:B-instance}\fi}
			\ifnum\mode=0
				\renewcommand{\arraystretch}{0.97}
			\else
				\renewcommand{\arraystretch}{1.02}
			\fi
			\begin{figure}[tbp]
				\foreach\a in {0,1}{%
						\centering
						\begin{subfigure}{\textwidth}
							\ifnum\mode=0
								\small
							\fi
							\centering
							\begin{tabular}{p{14em}m{16em}}
								\toprule
								\csvreader[no head, late after line= \\]{%
									\dir/\index{\a}.csv
								}{}{%
								\ifnum\thecsvrow=16 \midrule\fi\csvcoli & \csvcolii
								}
								\bottomrule
							\end{tabular}
							\caption{Instance~\AB{\a}}\mylabel{\a}
							\vspace{15pt}
						\end{subfigure}
					}
				\vspace{-15pt}
				\caption[Two instances sampled from recidivism dataset]{%
					Two instances sampled from recidivism dataset.
				}\label{fig:instance}
			\end{figure}
		}
		{%
			\ifnum\mode=0
				\def\scale{0.32}
				\def\imgwidth{0.45\textwidth}
				\def\hspacebase{\hspace{-0.5em}}
			\else
				\def\scale{0.315}
				\def\imgwidth{0.495\textwidth}
				\def\hspacebase{\hspace{-1.5em}}
			\fi
			\def\vspacebase{\vspace{0.5em}}
			\def\vspacebeforecaption{\vspace{-0.4em}}
			\begin{figure}[p]
				\centering
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{1.0em}
					\includegraphics[scale=\scale]{\dir/lime-\Asample}
					\vspacebeforecaption
					\caption{LIME}\label{fig:A-lime}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{0.5em}
					\includegraphics[scale=\scale]{\dir/newlime-\Asample-70}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.70$)}\label{fig:A-rlime-70}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/newlime-\Asample-80}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.80$)}\label{fig:A-rlime-80}
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{-0.5em}
					\includegraphics[scale=\scale]{\dir/newlime-\Asample-90}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.90$)}\label{fig:A-rlime-90}
				\end{subfigure}
				\caption[Explanation for Instance A by LIME and R-LIME]{%
					Explanation for Instance A by LIME and R-LIME.
				}\label{fig:A}
			\end{figure}
			\begin{figure}[p]
				\centering
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{-0.7em}
					\includegraphics[scale=\scale]{\dir/lime-\Bsample}
					\vspacebeforecaption
					\caption{LIME}\label{fig:B-lime}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{0.5em}
					\includegraphics[scale=\scale]{\dir/newlime-\Bsample-70}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.70$)}\label{fig:B-rlime-70}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{-0.8em}
					\includegraphics[scale=\scale]{\dir/newlime-\Bsample-80}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.80$)}\label{fig:B-rlime-80}
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{-0.8em}
					\includegraphics[scale=\scale]{\dir/newlime-\Bsample-90}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.90$)}\label{fig:B-rlime-90}
				\end{subfigure}
				\caption[Explanation for Instance B by LIME and R-LIME]{%
					Explanation for Instance B by LIME and R-LIME.
				}\label{fig:B}
			\end{figure}
		}
	}
\subsubsection{Experimental Results}
The results of the experiment are shown in \cref{fig:A,fig:B}.
The values assigned to each attribute name represent the contribution
(weights of the learned linear classifier)
to the output of the black-box classifier,
normalized such that the absolute sum is 1.
The figures also display the 5 attributes with the highest absolute contribution.
